Benchmarking neuromorphic systems with Nengo

### 2015-09-14 paper/rebuttals.txt

> This work addresses a very challenging topic - that is, how to
> objectively compare systems designed for different objectives in a
> high-dimensional benchmarking space. The approach of choosing
> cognitive tasks and measuring the ability of neuromorphic systems to
> perform them is insightful and relevant.

> It does, however, represent only an initial step towards obtaining
> the necessary concrete metrics to measure neuromorphic systems. Many
> quantities remain untested in this framework; researchers who build
> neuromorphic systems designed for neural simulation, the target
> audience of this manuscript, focus on optimizing many of the
> following quantities: simulation speed, simulation accuracy, energy
> per synaptic event, accurate timing, number of neurons represented,
> number of connections represented, repertoire of neural models,
> online learning, and likely others. Of these, the test framework
> objectively compares only along the first two, though the compliance
> tests likely would stress some of the others. I hope to see future
> work along these lines to expand the reach of these tests.

> Nonetheless, this work provides a substantive proposal to begin
> evaluating neuromorphic neural simulators.

Thank you for your feedback and helpful suggestions!
This is definitely just an initial step,
and we hope that we will be able to collect benchmark models
from many different researchers.

@@

  One inherent weakness of using Nengo as a standard platform for
  benchmarking neuromorphic systems is that new benchmarking
  capabilities may take a long time to be standardized and
  developed. If one wishes to add a new metric, such as power
  consumption (as was done in Stromatias et al., 2013), we must first
  come to a consensus on a suitable interface to this information
  through Nengo. Once consensus is reached, it must be implemented and
  tested on all backends before benchmarks can be written using that
  quantity. Despite this limitation, we believe that it is possible to
  use Nengo to collect power consumption information, and plan to
  implement energy efficiency comparisons in future work.

4. Other comments on the materials and methods.
-----------------------------------------------

> I found the separation in Background and Methods helpful for
> understanding, even if it does not follow the standard article
> structure. Some points should be clarified:
>
> - The benchmarks make use of current input to neurons. However,
>   quite a lot of neuromorphic systems do only offer spike input, or
>   current inputs are not so flexible to program. Of course, current
>   input can be emulated by spike input, but that may add some
>   additional noise/inaccuracy. This should be dicsussed in the
>   Background/Methods sections; also, an indication on how the
>   currents look like in the benchmark models would be helpful, so
>   that designers of neuromorphic systems get an idea of what would
>   be required.

This is an excellent point that we really should have addressed
in the initial submission. The paper that we cited in section
on the SpiNNaker backend (Mundy et al., 2015) gives a detailed
account of how SpiNNaker uses decoded values to change the
input currents to individual neurons, and two prior publications
(Galluppi et al., 2012 and Choudhary et al., 2012) describe processes
to implement Nengo with primarily spike-based communication
on SpiNNaker and Neurogrid, respectively.
We have added a paragraph to the Methods section
with brief descriptions of how SpiNNaker handles currents,
and how other neuromorphic hardware can be made to handle
currents from Nengo as well in lines 285-295.
The paragraph reads:

@@

  The only parameter on the network is an optional integer seed;
  setting this should make all of the random factors in the model
  deterministic, which is important for testing and debugging models,
  but is not used for benchmarking.

> - The authors state that Nengo is well-suited for generating large
>   networks. However, the tests they present are only limited in
>   size. A comment on larger benchmark models in the test suite
>   and/or the possibility to scale up the present benchmarks would be
>   desirable.

We have added an explicit mention of the size of Spaun on lines 400-401
(since Spaun is built with Nengo and shows that it is possible
to generate large networks). We have also added a short paragraph
to the Test Models subsection discussing larger models;
see lines 255-259 which read:

  We have used relatively small models run for short times in order to
  run many iterations on all backends.  However, all of these models
  could be made significantly larger by increasing the number of
  neurons used (i.e., adjusting the n_neurons parameter on ensembles),
  and increasing the dimensionality of the signals represented in the
  model.  Additionally, existing tests of the semantic pointer
  architecture within Nengo use significantly more neurons than the
  models presented here, and could be adapted into benchmarks.

@@

7. Please add here any further comments on this manuscript.
-----------------------------------------------------------

> Overall, the article presents an attractive idea for benchmarking
> neuromorphic systems. In my opinion its main weakness is the lack of
> embedding in the current literature. Apart from that, it is a nice
> contribution to the field, and the open source code is a good
> opportunity for other neuromorphic system operators for benchmarking
> their systems.
>
> Minor:
>
> - missing citation details for Stewart et al. 2010

Corrected, thank you! In correcting this, we noticed that
Mundy et al., 2015 was also missing details,
which have been filled in.

Also thank you for all of your feedback!
We feel that the manuscript is now much stronger
thanks to the additions in response to your feedback.

### 2015-11-04 paper/paper.tex
Make typographic corrections from typesetter

An example script showing how Nengo can
express a functional network
with 500~neurons and 6~connections
(representing 50,000~connection weights
and 100~direct current injection sites)
in under 20~lines of code
can be seen in Appendix~A.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework.\footnote{Available at \url{http://pytest.org/}.}
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
A demonstration of how these fixtures
are used in test scripts
can be seen in Appendix~A.

@@

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two or more runs of the same test.
The runs might represent multiple versions
of the same backend---one before a
speedup and one after---or multiple
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Appendix~A for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

\paragraph{Compliance}
is the number of tests that a particular backend
passes successfully, relative to the reference
backend, which must pass all tests.
While this metric can give an indication
of how many features a particular backend supports,
it does not take into account how commonly used
a particular feature is;
for this reason, a backend that has relatively
low compliance can still be useful
in many situations.

\paragraph{Accuracy}
measures how well a particular backend
implements a desired model.
The actual accuracy metric depends
on the desired model,
but in simple cases can be
the root mean squared error (RMSE)
between a desired signal
and the actual signal decoded
from an ensemble.

\paragraph{Speed}
measures the amount of time it takes
for a backend to run a model.
Primarily, we are concerned with
the amount of time taken
in \texttt{run} calls;
however, we also separately measure how long
each backend spends building each model
(i.e., the time spent in \texttt{\_\_init\_\_}),
as models must be built before being run.

@@

Each test has several parameters
that can be varied
in order to benchmark a wide variety of cases
(see Appendix~A for an example).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{https://github.com/ctn-archive/bekolay-fnme2015}
for those who wish to explore additional cases.

\paragraph{1. Communication channel chain.}
In this model,
five ensembles,
comprised of 100 leaky integrate-and-fire (LIF) neurons each
(500 total),
are connected in series,
with a communication channel
(i.e., the identity function)
computed across each connection.
This model, therefore, attempts to faithfully
communicate an input signal to
the last ensemble in the chain.
The input signal used is a static vector.
This model tests how
robust each backend is to the noise introduced
by representing a vector space
in the activity of spiking neurons.

\paragraph{2. Two-dimensional product.}
In this model,
the scalar product is computed
from a two-dimensional ensemble
comprised of 100 LIF neurons.
Since the product is nonlinear,
we use a space-filling curve
(the Hilbert curve; \citealp{hilbert1891})
as the input signal to ensure that
we sample the entire two-dimensional space.
This model tests each backend's
ability to compute a nonlinear function,
albeit a low-dimensional one.
Despite the low dimensionality,
two-dimensional products are frequently used
in large-scale models;
it is the primary operation needed to support
symbol-like processing using neuron-like elements,
for example \citep{eliasmith2013}.

\paragraph{3. Controlled oscillator.}
In this model,
a three-dimensional ensemble is recurrently connected
such that the first two dimensions continuously
traverse a limit cycle
(i.e., they implement a cyclic attractor),
and the third dimension controls the speed
and direction of the oscillation.
As input, we provide an initial stimulus
to start the oscillation,
and provide a control signal
to cause oscillation at
2~Hz, 1~Hz, 0~Hz, -1~Hz, and -2~Hz.
Negative frequencies indicate oscillations
in the reverse direction.\footnote{There is some evidence
  that neural oscillators such as locomotive central pattern generators
  accomplish forward and backward locomotion
  with the same neural mechanism \citep{duysens1998}.
  While this does not mean that all neural oscillators must
  be able to operate in the reverse direction,
  we believe that the capability to traverse
  the limit cycle forward and backward is advantageous,
  and in the current design requires no additional neural resources.}
In total, the model is comprised of 600 LIF neurons.
This model tests each backend's ability
to stably implement a dynamical system,
which is required for many cognitive functions
like working memory and motor control.

\paragraph{4. Basal ganglia sequence.}
In this model, brain structures
known collectively as the basal ganglia
are constructed from 4900 LIF neurons,
and are organized such that they iterate
through a repeating set of actions.
This model has been used to investigate
action selection and learning \citep{stewart2012a}
and the switching time between actions
has been closely mapped to human decision making
\citep{stewart2010}.
This benchmark tests the ability
to construct this model,
and evaluates the time needed
to transition between actions.

@@

\paragraph{1. Reference (\texttt{nengo}).}
The reference backend is designed to run quickly
on any general purpose computer
by using NumPy \citep{vanderwalt2011}
for fast vectorized computations.
It is included with the Nengo frontend
as \texttt{nengo.Simulator}.
The reference backend offers
the most features, but does not target
specialized hardware.

\paragraph{2. Distilled (\texttt{nengo\_distilled}).}
The distilled backend is intended as a teaching tool,
and as a template for building new backends.
It can also run on any general purpose computer,
and also uses NumPy for fast vectorized computations,
but does not aim to implement all of the features
implemented by the reference backend,
and omits some optimizations that obfuscate code.
Therefore, the distilled backend is easier to read
and suitable for learning about Nengo,
but is expected to be slower in terms of run time.

\paragraph{3. OpenCL (\texttt{nengo\_ocl}).}
The OpenCL backend uses the Open Computing Language
(OpenCL; \citealp{stone2010})
to run Nengo models on many different computing devices,
including graphical processing units (GPUs),
and field-programmable gate arrays (FPGAs).
In contrast to the distilled backend,
it is designed to run Nengo models
as quickly as possible,
using fast general purpose computing devices
like GPUs and any optimizations available,
at the cost of code readability.

\paragraph{4. Brainstorm software (\texttt{nengo\_brainstorm}).}
The Brainstorm backend is a software implementation
of a new neuromorphic chip
based partly on Neurogrid \citep{benjamin2014}
currently in development
by the Brains in Silicon lab
at Stanford University.\footnote{
  Some details available at \url{http://brainstorm.stanford.edu/projects/}.}
The software backend does not aim for speed;
instead, it attempts to emulate the proposed hardware
in order to test its applicability
for large-scale neural models.
If the emulated hardware can perform well,
then it follows that actual hardware
will also perform well,
but will be much faster.

\paragraph{5. SpiNNaker hardware (\texttt{nengo\_spinnaker}).}
The SpiNNaker backend \citep{mundy2015}
targets the eponymous neuromorphic hardware
developed by \citet{furber2014}.
In contrast to the Brainstorm backend,
this backend targets physical neuromorphic hardware,
and therefore is concerned
both with accuracy and speed.

@@

\section*{Appendix A: Chained Communication Channel model}

@@

\begin{figure}[!ht]
\begin{center}
  \includegraphics[width=85mm]{fig6}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:06} Figure \arabic{figure}.}
       {Run speed for each model on each backend. Each bar represents
         the mean run speed across 50 instances of each model;
         error bars are bootstrapped 95\% confidence intervals.
         Run speed is measured as the simulation time plus overhead,
         relative to real time; e.g., a value of two indicates that
         the calling \texttt{run} for a number of seconds takes
         the model two times that number of real seconds to complete the
         \texttt{run} call.
         BG Sequence * refers to the basal ganglia sequence model
         in which passthrough nodes have been removed.}

>>>>

An example script showing how Nengo can
express a functional network
with 500~neurons and 6~connections
(representing 50,000~connection weights
and 100~direct current injection sites)
in under 20~lines of code
can be seen in the Appendix.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework.\footnote{Available at \url{http://pytest.org/}.}
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
A demonstration of how these fixtures
are used in test scripts
can be seen in the Appendix.

@@

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two or more runs of the same test.
The runs might represent multiple versions
of the same backend---one before a
speedup and one after---or multiple
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See the Appendix for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

\subsubsection{Compliance}

Compliance is the number of tests that a particular backend
passes successfully, relative to the reference
backend, which must pass all tests.
While this metric can give an indication
of how many features a particular backend supports,
it does not take into account how commonly used
a particular feature is;
for this reason, a backend that has relatively
low compliance can still be useful
in many situations.

\subsubsection{Accuracy}

Accuracy measures how well a particular backend
implements a desired model.
The actual accuracy metric depends
on the desired model,
but in simple cases can be
the root mean squared error (RMSE)
between a desired signal
and the actual signal decoded
from an ensemble.

\subsubsection{Speed}

Speed measures the amount of time it takes
for a backend to run a model.
Primarily, we are concerned with
the amount of time taken
in \texttt{run} calls;
however, we also separately measure how long
each backend spends building each model
(i.e., the time spent in \texttt{\_\_init\_\_}),
as models must be built before being run.

@@

Each test has several parameters
that can be varied
in order to benchmark a wide variety of cases
(see the Appendix for an example).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{https://github.com/ctn-archive/bekolay-fnme2015}
for those who wish to explore additional cases.

\subsubsection{Communication channel chain.}

In this model,
five ensembles,
comprised of 100 leaky integrate-and-fire (LIF) neurons each
(500 total),
are connected in series,
with a communication channel
(i.e., the identity function)
computed across each connection.
This model, therefore, attempts to faithfully
communicate an input signal to
the last ensemble in the chain.
The input signal used is a static vector.
This model tests how
robust each backend is to the noise introduced
by representing a vector space
in the activity of spiking neurons.

\subsubsection{Two-dimensional product.}

In this model,
the scalar product is computed
from a two-dimensional ensemble
comprised of 100 LIF neurons.
Since the product is nonlinear,
we use a space-filling curve
(the Hilbert curve; \citealp{hilbert1891})
as the input signal to ensure that
we sample the entire two-dimensional space.
This model tests each backend's
ability to compute a nonlinear function,
albeit a low-dimensional one.
Despite the low dimensionality,
two-dimensional products are frequently used
in large-scale models;
it is the primary operation needed to support
symbol-like processing using neuron-like elements,
for example \citep{eliasmith2013}.

\subsubsection{Controlled oscillator.}

In this model,
a three-dimensional ensemble is recurrently connected
such that the first two dimensions continuously
traverse a limit cycle
(i.e., they implement a cyclic attractor),
and the third dimension controls the speed
and direction of the oscillation.
As input, we provide an initial stimulus
to start the oscillation,
and provide a control signal
to cause oscillation at
2~Hz, 1~Hz, 0~Hz, -1~Hz, and -2~Hz.
Negative frequencies indicate oscillations
in the reverse direction.\footnote{There is some evidence
  that neural oscillators such as locomotive central pattern generators
  accomplish forward and backward locomotion
  with the same neural mechanism \citep{duysens1998}.
  While this does not mean that all neural oscillators must
  be able to operate in the reverse direction,
  we believe that the capability to traverse
  the limit cycle forward and backward is advantageous,
  and in the current design requires no additional neural resources.}
In total, the model is comprised of 600 LIF neurons.
This model tests each backend's ability
to stably implement a dynamical system,
which is required for many cognitive functions
like working memory and motor control.

\subsubsection{Basal ganglia sequence.}

In this model, brain structures
known collectively as the basal ganglia
are constructed from 4900 LIF neurons,
and are organized such that they iterate
through a repeating set of actions.
This model has been used to investigate
action selection and learning \citep{stewart2012a}
and the switching time between actions
has been closely mapped to human decision making
\citep{stewart2010}.
This benchmark tests the ability
to construct this model,
and evaluates the time needed
to transition between actions.

@@

\subsubsection{Reference (\texttt{nengo}).}

The reference backend is designed to run quickly
on any general purpose computer
by using NumPy \citep{vanderwalt2011}
for fast vectorized computations.
It is included with the Nengo frontend
as \texttt{nengo.Simulator}.
The reference backend offers
the most features, but does not target
specialized hardware.

\subsubsection{Distilled (\texttt{nengo\_distilled}).}

The distilled backend is intended as a teaching tool,
and as a template for building new backends.
It can also run on any general purpose computer,
and also uses NumPy for fast vectorized computations,
but does not aim to implement all of the features
implemented by the reference backend,
and omits some optimizations that obfuscate code.
Therefore, the distilled backend is easier to read
and suitable for learning about Nengo,
but is expected to be slower in terms of run time.

\subsubsection{OpenCL (\texttt{nengo\_ocl}).}

The OpenCL backend uses the Open Computing Language
(OpenCL; \citealp{stone2010})
to run Nengo models on many different computing devices,
including graphical processing units (GPUs),
and field-programmable gate arrays (FPGAs).
In contrast to the distilled backend,
it is designed to run Nengo models
as quickly as possible,
using fast general purpose computing devices
like GPUs and any optimizations available,
at the cost of code readability.

\subsubsection{Brainstorm software (\texttt{nengo\_brainstorm}).}

The Brainstorm backend is a software implementation
of a new neuromorphic chip
based partly on Neurogrid \citep{benjamin2014}
currently in development
by the Brains in Silicon lab
at Stanford University.\footnote{
  Some details available at \url{http://brainstorm.stanford.edu/projects/}.}
The software backend does not aim for speed;
instead, it attempts to emulate the proposed hardware
in order to test its applicability
for large-scale neural models.
If the emulated hardware can perform well,
then it follows that actual hardware
will also perform well,
but will be much faster.

\subsubsection{SpiNNaker hardware (\texttt{nengo\_spinnaker}).}

The SpiNNaker backend \citep{mundy2015}
targets the eponymous neuromorphic hardware
developed by \citet{furber2014}.
In contrast to the Brainstorm backend,
this backend targets physical neuromorphic hardware,
and therefore is concerned
both with accuracy and speed.

@@

\section*{Appendix: Chained Communication Channel model}

@@

\begin{figure}[!ht]
\begin{center}
  \includegraphics[width=85mm]{fig6}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:06} Figure \arabic{figure}.}
       {Run speed for each model on each backend. Each bar represents
         the mean run speed across 50 instances of each model;
         error bars are bootstrapped 95\% confidence intervals.
         Run speed is measured as the simulation time plus overhead,
         relative to real time; e.g., a value of two indicates that
         calling \texttt{run} for a number of seconds takes
         the model two times that number of real seconds to complete the
         \texttt{run} call.
         BG Sequence * refers to the basal ganglia sequence model
         in which passthrough nodes have been removed.}
